---
title: Homework 2
description: |
  Homework 2 - Reading and pulling data.
author: Molly Hackbarth
date: 09-16-2021
output:
  distill::distill_article:
    self_contained: no
draft: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Reading Data into R

You can read data into R through a couple of methods. The first method is to use the library(datasets) function which will allow you to call dataset from the library. This is often seen using the dataset *iris* in examples. If you call table(iris) it will load most of the iris dataset.

>library(datasets)<br>
>table(iris)

## Reading Data into R through your own datasets

Reading your own dataset into R is a different process. In order to do this you must pull it from your **working directory.** To find your this directory you can use the getwd() function. *If you're confused about what a working directory is, it is a directory is your computer file folders (i.e. Downloads, Documents, etc.), and when R is running it's working in one of these folders. Hence the name **working directory**.*

>library(tidyverse)<br>
>library(readxl)<br>
>StateCounty2012 <- read_excel("/_data/StateCounty2012.xls")<br>
View(StateCounty2012)

## You can also use HERE

You can also use HERE to make links easier. *you can read more about HERE here: https://github.com/jennybc/here_here#* One reason to use HERE is it allows you to bypass the issue of setwd(), allowing you to change your working directory file, which can cause issues! **A relative path to the project root directory will always be created using here().**

>library(here)
>library(tidyverse) <br>
>library(readxl)<br>
>StateCounty2012 <- read_excel(here("_data", "StateCounty2012.xls"))<br>
>View(StateCounty2012)'

### Notes
- If you are having trouble reading a file make sure to check knit and then go to the knit directory and check Project Directory!
- if you are having trouble viewing the worksheet when you are knitting it on a mac you may need to download xquartz found here: https://www.xquartz.org/
- To render a file you may need to use ../_data/StateCounty2012.xls however when you submit it to github you'll want to remove the periods to show it as /_data/StateCounty2012.xls so that will work on the instructor's computer. 
- If you have the more subfolders to get to your file in HERE, you would add them before the worksheet/image/project. (i.e. here("images", "best-dogs", "goodestboy.jpg"))
- A great rundown on how HERE works is here: http://jenrichmond.rbind.io/post/how-to-use-the-here-package/

## Preview the data

### An example of untidy data

Using head() you can preview the data. You will notice that it's not very tidy.

>library(here)<br>
library(readxl)<br>
StateCounty2012 <- read_excel(here("_data", "StateCounty2012.xls"))<br>
head(StateCounty2012)

```{r print}
library(here)
library(readxl)
StateCounty2012 <- read_excel(here("_data", "StateCounty2012.xls"))
head(StateCounty2012)
```

### An example of tidy data

Here is an example of tidy data.

>library(here)<br>
library(readr)<br>
Eggs <- read_csv(here("_data", "eggs_tidy.csv"))<br>
head(Eggs)

```{r tidy data}
library(here)
library(readr)
library(knitr)
Eggs <- read_csv(here("_data", "eggs_tidy.csv"))
head(Eggs)
```

### Tibble

You can also use tibble directly, which is part of tidyverse, to create a table for your data.

>library(tidyverse)<br>
library(here)<br>
Eggs <- read_csv(here("_data", "eggs_tidy.csv"))<br>
as_tibble(Eggs)

```{r tibble}
library(tidyverse)
library(here)
Eggs <- read_csv(here("_data", "eggs_tidy.csv"))
as_tibble(Eggs)
```

## Using Kable

If you would like to show the full data table you can using kable. Below you will see the kable version for StateCounty2012 and Eggs. 

>kable(Eggs, caption = "Here is the tidy data of Eggs")<br>
kable(StateCounty2012, caption = "Here is the untidy data of StateCounty2012")

```{r kable}
kable(Eggs [1:4, ], caption = "Here is the tidy data of Eggs")
kable(StateCounty2012 [1:4, ], caption = "Here is the untidy data of StateCounty2012")
```

## Using rmarkdown

You can use rmarkdown for paged tables. 

>library(rmarkdown)<br>
paged_table(Eggs)<br.
paged_table(StateCounty2012)

```{r rmarkdown}
library(rmarkdown)
paged_table(Eggs)
paged_table(StateCounty2012)
```

## Editing a file 

To edit data you can install editData (install.packages("editData")). You can read more about it here: https://cran.r-project.org/web/packages/editData/vignettes/editData.html

>require(editData)<br>
>tibble(StateCounty2012)<br>
>result <- editData(StateCounty2012)


Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.


