---
title: HW#2 - Getting familiar with the US public schools data set
description: |
  Using the public school characteristic dataset collected in the 2017-2018 school year
author: Brittany Kenison
date: 09-16-2021
output:
  distill::distill_article:
    self_contained: no
draft: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
```

## Introduction

Let's get familiar with the data set by looking at the a few records:


```{r publicschool}
#read data from public school characteristics file
psc <- read_csv("../../_data/Public_School_Characteristics_2017-18.csv")

#display the first 6 rows of data
head(psc)
```

## Rows and Columns
Hmm.  That looks like a lot of data.  How many rows and columns are here?

```{r dimensions}
dim(psc)
```

## Column Names

Finally, the column names may help us understand the types of information collected. 

```{r columnnames}
colnames(psc)
```

## The End
Now that we are more familiar with the date set, the next blog post will start to wrangle the data for our eventual analysis. 
