---
title: "HW3_Updated_Justin Flattery"
author: "Justin"
date: "10/5/2021"
output: 
  distill::distill_article:
    self_contained: false
draft: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(blogbuilder)
library(distill)
library(tidyverse)
```

### I am looking to first read in a dataset. Specifically the dataset for debt_in_trillions. 

```{r , message=FALSE}
debt<- readxl::read_excel("_data/debt_in_trillions.xlsx")
```

### I will now use the select function to only view Mortgage, HE Revolving, Student Loan, Credit Card, Total

```{r, message = false}
select(debt, "Mortgage", "HE Revolving", "Auto Loan", "Credit Card", "Student Loan", "Total")
```

#### Now that I have the columns I want, I will now narrow down the rows of interest, specifically I want to look at when Mortgage's are greater then 8. To do this I will use the filter function

```{r, message=false}
filter(debt,Mortgage > 8)
```
### Next I will use the arrange to sort the columns by lowest to highest Student Loan %

```{r, message = false}
arrange(debt, desc('Student Loan'))
```
### Finally, I will use the summarize function
```{r, message = false}
summarise(debt)
```
### It seems the summarize function does not produce any compelling results because there is no function grouping these together
