---
title: First blog post using Distill
description: I'm going to make my first blog post. Once I figure that out, I'll try
  reading in some data.I am also experimenting with R Markdown text formatting such
  as **bold text** and *italics*
author: ~
name: Amy Franks
date: 09-29-2021
output:
  distill::distill_article:
    self_contained: no
draft: yes
---

### First I need to load my packages by using the **library** command and this chunk:

{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(readxl)  

### But oh no. I want this code to be displayed in 4 rows. This can be done by adding *two spaces* at the end of each line:

{r setup, include=FALSE}  
knitr::opts_chunk$set(echo = TRUE)  
library(tidyverse)  
library(readr)  
library(readxl)  
library(csv)  


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(readxl)
library(csv)
```

### I'm going to in work with data from the set about cereal. I want to first *establish my working directory*. I can find the current working directory with the command *getwd()*

```{r get working directory}
getwd()
```

### I'm going to set the directory just to be sure I'm working with the right folders and files.

```{r set working directory}
#setwd("~/R/DACSS601Fall21/_posts/2021-09-29-first-blog-post-using-distill")
```

### Now I'm going to *read in a csv data set about cereal*. This following chunk creates a new tab that contains a table with the cereal data

```{r read in cereal info}
cereal <- read.csv("cereal.csv")
View(cereal)
```

### If I need to figure out the dimensions of the data set, or get other information, I can use a number of commands. Using *dim()* returns info on the dimensions. In this case, there are 20 rows and 4 columns in the *cereal.csv* file.

```{r get some info about the scope of the cereal dataset} 
dim(cereal)
```

### I can use *colnames()* to get the name of each column. In this case, the column names are ***"Cereal", "Sodium", "Sugar" and "Type"***

```{r}
colnames(cereal)
```

### Since I assigned the csv file to the variable *cereal*, I can retrieve the data by simply calling the function by writing *"cereal"*.

```{r}
cereal
```

### These are the very basics of reading in data. I am excited to *keep learning* how to use R and R Markdown, and how to manipulate and utilize data.





Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.


