---
title: Quinn He HW2
description: "Homework 2 where I try to open datasets and see what I can do with it.
  \n"
author: Quinn He
date: 09-16-2021
output:
  distill::distill_article:
    self_contained: no
draft: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


```{r}
hotel_bookings <- read_csv("../../_data/hotel_bookings.csv")


```

## Amount of weekend nights stayed in (per party)

```{r}

ggplot(hotel_bookings, aes(stays_in_weekend_nights)) + geom_bar()

```
## Guests that have children

```{r}
ggplot(hotel_bookings, aes(children)) + geom_bar()
```

# I wanted to see if there is a relationship between people who stay in on weekend nights and those who have children. It does not appear a large number of guests have children so what else could it be? Perhaps babies?

```{r}
ggplot(hotel_bookings, aes(babies)) + geom_bar()
```
# Cleary that also does not have an impact on the decision.

# Months arrived by customeres
```{r}
 ggplot(data = hotel_bookings) +
geom_bar(mapping = aes(x = arrival_date_month))
```


# Most popular hotel type
```{r}
 ggplot(data = hotel_bookings) + 
  geom_bar(mapping = aes(x = hotel))
```


