---
title: HW3 Data Wrangling
description: |
  Lets get this data wrangled!
author: Eliza Geeslin
date: 10-04-2021
output:
  distill::distill_article:
    self_contained: no
draft: yes
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(stringr)
library(rmarkdown)
library(knitr)
library(formattable)
library(vctrs)
```

## Step 1: Read in the Data

The first step is getting everything up and running. 
* This dataset has a few tabs, so even though I am using the first tab (the default) I want to ensure we are using the correctly named tab every time. 
* There are actually two data sets (eggs and pounltry). I am going to be honest - I did not totally realize this until halfway through the wrangingling, so I decided to try to tidy the dataset into one that could be filtered down to either chicken or eggs. Not totally sure it worked :). 

```{r}
organic_egg_poultry <- read_excel("../../_data/organiceggpoultry.xls", sheet = "Data") # assign dataset to variable

#head(organic_egg_poultry)

colnames(organic_egg_poultry) # colnames() returns column names
```
## Step 2: Tidy the Data

I am going to do a few things to tidy the data:
* Get rid of extra columns and rows
* Adjust column names (for future pivoting)
* Adjust data types

First, the column names are off, so I am going to skip 4 rows to get the actual column names. 

```{r}
organic_egg_poultry <- read_excel("../../_data/organiceggpoultry.xls", sheet = "Data", skip = 4)

paged_table(organic_egg_poultry)
```

Some of the column names are a little warped, so I am going to rename them: 
* I will indicate the ones to be deleted ("del")
* I am trying to name them in a way that will help later on - I think I might have to break out some of these columns because it seems like they hold multiple variables (Produce, Amount, ProduceType). This is where I realized there are really two distinct data sets, but c'est la vie!

```{r}
organic_egg_poultry <- read_excel("../../_data/organiceggpoultry.xls", sheet = "Data", skip = 5, col_names = c("Month", "ExtraLargeEggs_Dozen_Eggs", "ExtraLargeEggs_HalfDozen_Eggs", "LargeEggs_Dozen_Eggs", "LargeEggs_HalfDozen_Eggs", "del", "WholeChicken_Pound_Chicken", "BonelessChickenBreast_Pound_Chicken", "BoneChickenBreast_Pound_Chicken", "WholeChickenLegs_Pound_Chicken", "ChickenThighs_Pound_Chicken"))

paged_table(organic_egg_poultry)
```
Next, I noticed something weird with the month "February" in the dataset that will mess up the column splitting/filling (the next two steps), so I went ahead fixed that. 

```{r}
organic_egg_poultry2 <- organic_egg_poultry
organic_egg_poultry2[organic_egg_poultry2 == "February /1"] <- "February"

paged_table(organic_egg_poultry2)
```
Then, I split Month and Year.

```{r}
organic_egg_poultry3 <- organic_egg_poultry2 %>%
separate(col = 1, into=c('Month', 'Year'), sep = ' ')

paged_table(organic_egg_poultry3)
```

As we can see, the year doesn't fill up all the rows, so I used fill() to complete the rows.

```{r}
organic_egg_poultry3 <- organic_egg_poultry3 %>%
  fill(Year)

paged_table(organic_egg_poultry3)

```
Then I got rid of the extra column in between the Eggs and Chicken datasets (because I am still working with them together).

```{r}
organic_egg_poultry3 <- organic_egg_poultry3 %>%
  select(!starts_with("del"))

paged_table(organic_egg_poultry3)
```
## Step 3: Wrangle the Data

Before I pivot this data, I need to do something about the "too few" string in the chicken part of the dateset (I know this will cause an issue because I didn't do anything initially and I got an error: "`Can't combine `ExtraLargeEggs_Dozen_Eggs` <double> and `BoneChickenBreast_Pound_Chicken` <character>`").

I think the best bet here is to mutate "too few" into N/A? (although I am not totally sure if that will mess up the integrity of the data?). 

```{r}
organic_egg_poultry4 <- organic_egg_poultry3
organic_egg_poultry4[organic_egg_poultry4 == "too few"] <- NA

paged_table(organic_egg_poultry4)
```
I also have to make sure everything is the right data type. 

```{r}
organic_egg_poultry4 <- organic_egg_poultry4 %>%
  mutate(BoneChickenBreast_Pound_Chicken = as.double(BonelessChickenBreast_Pound_Chicken)) %>%
  mutate(ChickenThighs_Pound_Chicken = as.double(ChickenThighs_Pound_Chicken))

paged_table(organic_egg_poultry4)
```
At last I can pivot longer the data. This is where the column naming comes in handy!

```{r}
#pivot_long
organic_egg_poultry5 <- organic_egg_poultry4 %>%
  pivot_longer(
    cols = !Month & !Year, 
    names_to = c("Produce", "Amount", "ProduceType"), 
    names_sep = "_",
    values_to = "Price"
  )

paged_table(organic_egg_poultry5)
```

At this point the data is probably ready to arrange, filter, and summarize. I am going to arrange it by Produce Type (Eggs or Chicken). 

```{r}
organic_egg_poultry5 <- organic_egg_poultry5
arrange(organic_egg_poultry5, ProduceType, desc(Price))
```
Something that is a little annoying about this is that I wanted to arrange it so that the "Eggs" section was at the top, but it seems to only do alphabetical. 

Next, I am going to filter the data so we are only working with the Eggs data and then summarize.

```{r}
organic_egg_poultry6 <- organic_egg_poultry5
filter(organic_egg_poultry6, ProduceType == "Eggs") %>%
  group_by(Produce, Amount) %>%
  summarise(mean = mean(Price), max = max(Price), min = min(Price))
```



Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.


