---
title: 'Homework 3: Basic Data Wrangling'
description: |
  Practice wrangling data with the Australian Marraige Law Excel document.
author: Megan Georges
date: 10-05-2021
output:
  distill::distill_article:
    self_contained: no
draft: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(distill)
library(tidyverse)
library(dplyr)
library(stringr)
library(readxl)
```

# Read the data into R

* Specified that I wanted Table 2 from the Excel file 
* Skip the first 6 rows
* Provided preview of first 10 lines of data

```{r}
ausmarraige <- read_excel("../../_data/australian_marriage_law_postal_survey_2017_-_response_final.xls", "Table 2", skip=6) 
head(ausmarraige, 10)
```

# View column names

```{r}
colnames(ausmarraige)
```

# Select the columns that we want to keep and rename them

The essential columns in this dataset are the cities, response clear yes, response clear no, response not clear, and non responding. 

```{r}
ausmarraige1 <- select(ausmarraige, "...1", "no....2", "no....4", "no....11", "no....13")%>%
  rename(Cities=...1, Response_Clear_Yes=no....2, Response_Clear_No=no....4, Response_Not_Clear=no....11, Non_Response=no....13)%>%
  drop_na(Cities)

head(ausmarraige1)
```

# Remove division totals and footnotes

```{r}
ausmarraige1 <- ausmarraige1 %>%
  filter(!str_detect(Cities, "Total"))%>%
  filter(!str_starts(Cities, "\\("))%>%
  filter(Cities != "Australia")%>%
  filter(!str_starts(Cities, "\\©"))
```

# Create new column for the divisions

```{r}
ausmarraige1 <- ausmarraige1 %>%
  mutate(Divisions = case_when(
    str_ends(Cities, "Divisions") ~ Cities
  ))
ausmarraige2 <- ausmarraige1[, c("Divisions", "Cities", "Response_Clear_Yes", "Response_Clear_No", "Response_Not_Clear", "Non_Response")]
head(ausmarraige2)
```

# Summarise the tidier (but still needs work) data

```{r}
summary(ausmarraige2)
```

# Arrange data in descending order of clear Yes responses

```{r}
arrange(ausmarraige2, desc(Response_Clear_Yes))
```

