---
title: "Australia Data Stat"
author: "John C. Manyiel"
date: "9/29/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.






##Load packages

```{r}
#library(rprojroot)
library(readxl)
library(rmarkdown)
library(ggplot2)
library(rvest)
library(tibble)
library(lubridate)
library(tidyverse)
library(stringr)
library(pdftools)
library(xlsx)
library(dbplyr)
library(knitr)
theme_set(bayesplot::theme_default(base_family = "sans"))
```


##Read in the data set


```{r}

file_location <- file.path("/../Development",
                           "DatumAnalyticConsulting",
                           "DACSS601Fall21",
                           "_data",
                           "australian_marriage_law_postal_survey_2017_-_response_final.xls")
obs_count <-read_excel(
                path = file_location, 
                range = "Table 2!A8:P183", 
                col_names = c("area",
                              "Yes",
                              "Yes pct",
                              "No",
                                     "No pct",
                                     "Response Total",
                                     "Response Total pct",
                                     "blank",
                                     "Response clear",
                                     "Response clear pst",
                                     "Response not clear(b)",
                                     "Response not clear(b) pct",
                                     "Non-responding",
                                     "Non-responding pct",
                                     "Eligible Total",
                                     "Eligible Total pct"))


head(obs_count[1:6])

```


##Transform Data

Subset by Division and add Division column with row value "New South Wales"


```{r}


NewSouthWales<-  obs_count[2:48,] %>% as_tibble() %>% 
                                         mutate(Division = "New South Wales")

head(NewSouthWales[,17])


```



Subset by Division and add Division column with row value "Victoria"



```{r}




Victoria <-  obs_count[52:88,] %>% as_tibble() %>% 
                                         mutate(Division = "Victoria")

head(Victoria[,17])


```

Subset by Division and add Division column with row value "Queensland"


```{r}



Queensland <-  obs_count[92:121,] %>% as_tibble() %>% 
                                         mutate(Division = "Queensland")

head(Queensland[,17])


```

Subset by Division and add Division column with row value "South Australia"


```{r}



SouthAustralia <-  obs_count[125:135,] %>% as_tibble() %>% 
                                         mutate(Division = "South Australia")

head(SouthAustralia[,17])



```

Subset by Division and add Division column with row value "Western Australia"

```{r}



WesternAustralia <-  obs_count[139:154,] %>% as_tibble() %>% 
                                         mutate(Division = "Western Australia")

head(WesternAustralia[,17])

```

Subset by Division and add Division column with row value "Tasmania"


```{r}



Tasmania <-  obs_count[158:162,] %>% as_tibble() %>% 
                                         mutate(Division = "Tasmania")

head(Tasmania[,17])



```

Subset by Division and add Division column with row value "Northern Territory"

```{r}

NorthernTerritory <-  obs_count[166:167,] %>% as_tibble() %>% 
                                         mutate(Division = "Northern Territory")

head(NorthernTerritory[,17])

```


Subset by Division and add Division column with row value "Australian Capital Territory"


```{r}

CapitalTerritory <-  obs_count[171:172,] %>% as_tibble() %>% 
                                         mutate(Division = "Australian Capital Territory")


head(CapitalTerritory[,17])

```



Combine all Division variables


```{r}


obs_count01 <-  rbind( NewSouthWales, Victoria, Queensland, 
                       SouthAustralia, WesternAustralia, 
                       Tasmania, NorthernTerritory, 
                       CapitalTerritory) %>%  as_tibble()


dim(obs_count01)


```





Continue to transform the data set and remove blank field such as total and the blank value 


```{r}
obs_count02 <-  obs_count01 %>%
                    select(-blank)%>%
                          filter(!grepl("Total", area))%>%
                                  filter(!grepl("Total", area)) %>%drop_na()

head(obs_count02)



```

Graph Data

Begin  summarizing & visualizing your data


```{r}



```


```{r}

```


```{r}


```

