---
title: Homework 2- Reading in Data
description: |
  Reading in Data
author: Kirsten Johnson
date: 10-10-2021
output:
  distill::distill_article:
    self_contained: no
draft: yes
---
```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

### In this post, I will try reading in data. I will be reading in the Austrailian marriage tidy dataset. This data breaks down Austrailian marriage staticstics by territory.

#### Lets read in the australian marriage csv:

<!-- #read in data -->
```{r}
aus_marriage<-read_csv("../../_data/australian_marriage_tidy.csv")
```



#### To take a quick look at what the data contains, I want to view the csv column names.

<!-- #Get column names -->
```{r}
colnames(aus_marriage) 
```


#### Here are the column field definitions:
##### 1. "territory" lists the Australian territory name.
##### 2. "resp" denotes marital status as either 'yes', married or 'no', not married.
##### 3. "count" gives a count of how many people from a territory have a specific marital status.
##### 4. "percent" gives a percentage of the people from a territory with a specific marital status.
