---
title: "HW4: Univariate Statistics"
description: |
  2017 Australian Marraige Law dataset: tidying/cleaning the data, discussing the variables, incorporating some visualizations (all for practice)
author: Megan Georges
date: 10-30-2021
output:
  distill::distill_article:
    self_contained: false
draft: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(distill)
library(tidyverse)
library(dplyr)
library(stringr)
library(readxl)
library(ggplot2)
```

# Description of the Dataset and Variables

## 2017 Australian Marraige Law 

A postal survey was sent to Australians asking the question: Should the law be changed to allow same-sex couples to marry? The data was reported by the Australian Bureau of Statistics using Excel.

## Variables


* Divisions (Australian States/Territories)
* Cities (within the Divisions) 
* Response_Clear_Yes: those who clearly selected Yes in support of the new marraige law
* Response_Clear_No: those who clearly selected No to the new marraige law
* Response_Not_Clear: those in which researchers could not determine which answer the respondent intended to select
* Non_Response: those who did not return the survey or select any of the responses

***

# Tidying and Cleaning the Data

## Read the data into R

```{r}
ausmarraige <- read_excel("../../_data/australian_marriage_law_postal_survey_2017_-_response_final.xls", "Table 2", skip=6) 
head(ausmarraige, 10)
```

## View column names

```{r}
colnames(ausmarraige)
```

## Select the columns that we want to keep and rename them

```{r}
ausmarraige1 <- select(ausmarraige, "...1", "no....2", "no....4", "no....11", "no....13")%>%
  rename(Cities=...1, Response_Clear_Yes=no....2, Response_Clear_No=no....4, Response_Not_Clear=no....11, Non_Response=no....13)%>%
  drop_na(Cities)

head(ausmarraige1)
```

## Remove division totals and footnotes

```{r}
ausmarraige1 <- ausmarraige1 %>%
  filter(!str_detect(Cities, "Total"))%>%
  filter(!str_starts(Cities, "\\("))%>%
  filter(Cities != "Australia")%>%
  filter(!str_starts(Cities, "\\©"))
```

## Create new column for the divisions

```{r}
ausmarraige1 <- ausmarraige1 %>%
  mutate(Divisions = case_when(
    str_ends(Cities, "Divisions") ~ Cities
  ))
ausmarraige2 <- ausmarraige1[, c("Divisions", "Cities", "Response_Clear_Yes", "Response_Clear_No", "Response_Not_Clear", "Non_Response")]
head(ausmarraige2)
```

## Continue to fix the Divisions column using fill

```{r}
ausmarraige2 <- fill(ausmarraige2, Divisions, .direction = c("down"))
```

## Remove the Divisions from the Cities column using filter

```{r}
ausmarraige2 <- filter(ausmarraige2, !str_detect(Cities, "Divisions"))
head(ausmarraige2)
```

## Create a column to show Totals for each City 

```{r}
austotals <- ausmarraige2 %>%
  mutate(Total=rowsum(Response_Clear_Yes, Response_Clear_No, Response_Not_Clear, Non_Response))
head (austotals)
```

***

# Summary Descriptives of the Variables

## Divisions and Cities

Australia comprises of 6 states and 2 territories (thus, 8 Divisions). The number of Cities in each Division is displayed below:

```{r}
select(ausmarraige2, Divisions)%>%
  table()
```

**The total number of eligible participants, by division:**

```{r}
divtotals <- select(austotals, Divisions, Total) 
divtotals2 <- divtotals %>%
  pivot_wider(names_from = Divisions, values_from = Total, values_fn = sum)
head(divtotals2)
```

**A proportion table of the eligible participants based on Division:**

```{r}
prop.table(divtotals2)
```

## Participant Responses

**Descriptive statistics for Response_Clear_Yes:**

```{r}
summarize(ausmarraige2, Mean = mean(Response_Clear_Yes), Median = median(Response_Clear_Yes), Min = min(Response_Clear_Yes), Max = max(Response_Clear_Yes), SD = sd(Response_Clear_Yes), Var = var(Response_Clear_Yes), IQR = IQR(Response_Clear_Yes))
```

**Descriptive statistics for Response_Clear_No:**


```{r}
summarize(ausmarraige2, Mean = mean(Response_Clear_No), Median = median(Response_Clear_No), Min = min(Response_Clear_No), Max = max(Response_Clear_No), SD = sd(Response_Clear_No), Var = var(Response_Clear_No), IQR = IQR(Response_Clear_No))
```

**Descriptive statistics for Response_Not_Clear:**

```{r}
summarize(ausmarraige2, Mean = mean(Response_Not_Clear), Median = median(Response_Not_Clear), Min = min(Response_Not_Clear), Max = max(Response_Not_Clear), SD = sd(Response_Not_Clear), Var = var(Response_Not_Clear), IQR = IQR(Response_Not_Clear))
```

**Descriptive statistics for Non_Response:**

```{r}
summarize(ausmarraige2, Mean = mean(Non_Response), Median = median(Non_Response), Min = min(Non_Response), Max = max(Non_Response), SD = sd(Non_Response), Var = var(Non_Response), IQR = IQR(Non_Response))
```

# Experimenting with a Visualization

**Bar graph modeling the total number of each response:**

```{r}
responsetotals <- ausmarraige2 %>%
  select(Response_Clear_Yes, Response_Clear_No, Response_Not_Clear, Non_Response)%>%
  summarise(Yes=sum(Response_Clear_Yes), No=sum(Response_Clear_No), NotClear=sum(Response_Not_Clear), NonResponse=sum(Non_Response))
rownames(responsetotals) <- c("Total")

responsetotals <- pivot_longer(responsetotals, Yes:NonResponse, names_to = "Response", values_to = "Total")

ggplot(data = responsetotals) +
  geom_bar(mapping = aes(x = Response, y = Total), stat = "identity") +
  labs(title = "Should Australian Law Change to Allow Same-Sex Marraige?", subtitle = "2017 postal survey of Australians", x = "Type of Response", y = "Total # of Responses")
```






