---
title: Homework 1
description: |
  Homework 1 - basics of R.
author:
- name: Molly Hackbarth
date: 09-15-2021
output:
  distill::distill_article:
    self_contained: no
draft: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Learned from R

As a new student on R, I've learned some interesting new ideas!

### Basic Math

Below you will find how to do basic math **addition, subtraction, multiplication, and division.**   

```{r addition, subtraction, multiplcation, division}
5+4
3-2
4/1
3*4
5+4+3*14-2/2
```

### Logical Data

You can also use logic for data! This uses the **==, for equals, > greater than, < less than, and >= equal or greater or <= less than or equal to** signs!

```{r logic}
4==5
2==2
3>5
10>5
15>=15
1<=2
```


### Assigning Data 

You can also assign values to variables. 

```{r assigning data}

#create a variable x 

x <- 3

x+5

x*x

```
### Reading Data into R

You can also read and plot data into R! 

```{r readr}
library(datasets)
plot(iris)
```

### Summarazing Data 

You can also use head() to see the first few lines of a table.

```{r}
library(datasets)
head(iris)
```


### Vectoring Data

To vector data you must use **c() and <-**. Here is a quick set up of if you wanted to find the median of yards ran if it was multiplied by 100. 

```{r vector}
yards <- c(12,50,100,22)

run_yards <- yards*100
```

Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.


