---
title: 'Homework_2 '
description: Reading in Data
author:
- name: Cynthia Hester
date: 09-29-2021
output:
  distill::distill_article:
    self_contained: no
draft: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Reading in the first data set

Reading in or importing data files to RStudio is a necessary step to gain access to any files that are needed for cleaning or tidying. After imported data is cleaned, it is then more suitable for exploration.


As we know data formats are not homogeneous,and come in many different flavors. So,whether data is in **CSV, SPSS,XLSX,SAS,TXT,STATA,or HTML** as well as many other formats, there is usually R package to read in the data.


The first data set I will read in is from the included R package "Data Sets". It is the *MTCars* (MotorTrend) dataset which was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption as well as 10 aspects of automotive design and performance for 32 cars (1973-74).



**This R chunk loads in the data sets package and provides a summary of the statistics for the *mtcars* data set**

```{r}
library(datasets)
summary(mtcars)

```




**This R chunk uses an alternative to the *summary* function called *Skim*. *Skim* provides a comprehensive overview of the *mtcars* data set as well as providing a visualization of the  data in the rows represented by histograms.**


```{r}
library(skimr)
skim(mtcars)
```





**This R chunk exemplifies the granularity of the *Skim* package by selecting specific columns to summarize.**


```{r}
skim(mtcars,hp,wt)
```





**This R chunk provides the column names of the *mtcars* dataset using the colnames() function.**

```{r}
colnames(mtcars)
```





**This R chuck introduces the *dim()* function provides information on the dimensions of the data set,which shows this data array to have 32 rows and 11 columns.**

```{r}
dim(mtcars)
```




**This R chunk shows a generic visualization of the *mtcars* object using the *plot()* function.**
```{r}
plot(mtcars)
```



## The Second Data Set comes from the course csv file eggs_tidy. 

I wanted to try reading data in from an external data set, that used the csv format.


**This first R chunk reads in the eggs tidy csv data**

```{r}
library(readr)
eggs_tidy <- read_csv("../../_data/eggs_tidy.csv")
```


**Summarizes the eggs_tidy data set**
```{r}
summary(eggs_tidy)
```



**Summarizes data set using the skim function**
```{r}
skim(eggs_tidy)
```

**This chunk uses the *tibble function* which provides a more comprehensive and readable  data frame**

```{r}
library(tibble)
as_tibble(eggs_tidy)
```


